---
title: 'Statistical Inference: Basic inferential data analysis'
author: "Jacques Botes"
date: "September 2014"
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    keep_md: yes
---
```{r Load,echo=FALSE}
require(datasets)
require(ggplot2)
require(plyr)
      
data(ToothGrowth)
```

###Problem Statement
Load the ToothGrowth data and perform some basic exploratory data analyses 
Provide a basic summary of the data.
Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. 
(Use the techniques from class even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions. 

###About the data
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid). A data frame with 60 observations on 3 variables
```{r exploratory, echo=FALSE}
head(ToothGrowth)
summary(ToothGrowth)

##aggregate data
ToothGrowth.Summary <- ddply(ToothGrowth,.(dose,supp),summarise,len.avg=mean(len))
        
g <- ggplot(data=ToothGrowth, aes(x=factor(dose), y=len, fill=factor(supp))) +
        facet_grid(. ~ supp) +
       ## geom_bar(stat="identity") +
        geom_violin(col = "black", size = 2) +
        
        ylab('Tooth Length') + 
        xlab('dose (mg)') + 
        ggtitle('Tooth Length per supplement type and dosage (mg)')
g


g <- ggplot(data=ToothGrowth, aes(x = factor(supp), y = len, group = factor(dose))) +
        geom_line(size = 1, aes(colour = dose)) + 
        geom_point(size =10, pch = 21, fill = "salmon", alpha = .5)
g

```

###Summary of dataset: 
Interpreting the first graph we can see that the Orange Juice (OJ) supplement at lower doses results in a longer tooth length. With the 2mg dose the difference in tooth lengh is closely matched and seems to be working equally well.

###Testing

####Assumptions

* The variances are different for the populations

####Analysis
Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. 
```{r C_Supp}

t.test(len ~ supp, paired = F,var.equal=F, data = ToothGrowth)
```

We cannot reject the null hypothesis as there are a significant difference in tooth length betwen OJ and VC supplement types.

Lets look at it for specific dosages
```{r C_Dose}
tooth.D05 <- subset(ToothGrowth, dose == 0.5)
tooth.D1 <- subset(ToothGrowth, dose == 1.0)
tooth.D2 <- subset(ToothGrowth, dose == 2.0)

rbind(        
c('0.5mg conf interval',t.test(len ~ supp, paired = F, var.equal= F,data = tooth.D05)$conf),
c('1.0mg conf interval',t.test(len ~ supp, paired = F, var.equal= F,data = tooth.D1)$conf),
c('2.0mg conf interval',t.test(len ~ supp, paired = F, var.equal= F,data = tooth.D2)$conf)
)

```

###In Conclusion:
If we ignore the supplement types and just look at the difference in dosages we notice we can reject the null hypothesis for 0.5 and 1.0mg levels and not for the 2.0mg level. for the 2.0mg there is a significant difference in tooth length. 


Full markdown file:
