
---
title: 'Statistical Inference: Exponential Distribution Simulation'
author: "Jacques Botes"
date: "September 2014"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    theme: spacelab
  pdf_document: default
---

The exponential distribution can be simulated in R with `rexp(n, lambda)` where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. 
Set lambda = 0.2 for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
2. Show how variable it is and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.
4. Evaluate the coverage of the confidence interval for $1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$.
```{r Libraries, echo=FALSE, results='hold',warning=FALSE,}
require(ggplot2)
```
First setup the simulation and get the means of the 1000 simulations
```{r Setup_Simulation}
set.seed(9867)   ## to ensure reproducibility
n <- 1000        ## no of runs
sample.size <- 40 ## 40 samples in each run
lambda <- 0.2   ##variable input
dist <- matrix(rexp(sample.size*n, rate=lambda), ncol = sample.size, nrow=n)
dist.means <- rowMeans(dist) ##a  vector of n length with averages in each row based on sample size means
```

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
```{r Question 1}
hist(dist.means,breaks=sample.size,prob=T,col="green",main="Density of means",ylab="density")
abline(v=mean(dist.means),col="black",lwd=4)

center.ac <- mean(dist.means)
center.th <- 1/lambda
```
The Theoretical center of the distribution is calculated as $1/\lambda$ = 1/0.2 = `r center.th`. The center of the distribution is `r center.ac`

2. Show how variable it is and compare it to the theoretical variance of the distribution.

```{r}
sd.ac <- sd(dist.means)
sd.th <- (1/lambda)*(1/sqrt(sample.size))

var.ac <- sd.ac^2 ## = var(dist.means)
var.th <-((1/lambda)*(1/sqrt(sample.size)))^2
```
The Theoretical standard deviasion is $1/\lambda = * 1/\{\sqrt{n}}^2$.

xfit <- seq(min(dist.means), max(dist.means), length=100)

yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sample.size)))

##lines(xfit, yfit, pch=22, col="red", lty=2)


##v <- mean(dist.means) + c(-1,1)*1.96*sd(dist.means)/sqrt(sample.size)

lambda_vals <- seq(4, 6, by=0.01) ##we know the center is around 5
coverage <- sapply(lambda_vals, function(x) {
    mu.hats <- rowMeans(matrix(rexp(sample.size*n, rate=0.2),ncol = sample.size, nrow=n))
    err <- qnorm(0.975) * sqrt(1/lambda**2/sample.size)
    ll <- mu.hats - err
    ul <- mu.hats + err
    mean(ll < x & ul > x)
})

##ggplot(data.frame(lambda_vals, coverage), aes(x = lambda_vals, y = coverage)) + geom_line(size = 2) + ##geom_hline(yintercept = 0.95,col="red")+ ylim(.75, 1.0) + geom_vline(xintercept=v,col="red")

```
